---
title: "The left-right positions of British MPs inferred from a survey of local councillors"
subtitle: "Appendix"
published: "Working paper"
code-repo: "Access the code, data, and analysis at <>"
correspondence-prefix: "Correspondence concerning this article should be addressed to"
author:
  - name: Chris Hanretty
    email: chris.hanretty@rhul.ac.uk
    orcid: 0000-0002-3948-3914
    title: Professor of Politics
    affiliations: 
      - id: rhul
        name: "Royal Holloway, University of London"
        department: "Department of Politics, International Relations and Philosophy"
        address: McCrea 1-08
        city: Egham
        region: Surrey
        country: United Kingdom
        postal-code: TW20 0EX
    attributes:
      corresponding: true
  - name: Vasil Lazarov
    email: vasil.lazarov@survation.com
    title: PhD student
    affiliations:
      - id: survation
        name: Survation
      - ref: rhul
thanks: "Thanks to Survation for hosting these questions on their omnibus survey of local councillors. Particular thanks to David Izamoje for setting up the script seen by councillors. "
additional-info: |
  We have no known conflict of interest to disclose.
date: September 28, 2023
bibliography: biblio.bib
mainfont: "TeX Gyre Pagella"
sansfont: "TeX Gyre Heros"
prefer-html: true
toc: true
execute:
  echo: false
  message: false
  warning: false
format:
  hikmah-pdf: default
##  hikmah-manuscript-docx: default
---

```{r}
here::i_am("article/appendix.qmd")
suppressPackageStartupMessages(library(tidyverse))
library(knitr)
library(kableExtra)
library(stringr)
```

{{< pagebreak >}}

# Full listing of MPs

The table reports the positions and ranks of MPs in office at the time
of the survey. MPs appear with the party label under which they were
elected to the House of Commons, whether or not they retain the party
whip.

```{r}
plot_df <- readRDS(here::here("working", "partyonly_thetas.rds")) |>
    mutate(Party = factor(Party, ordered = TRUE),
           Party = reorder(Party, mean)) 

fit <- readRDS(here::here("working", "partyonly_fit.rds"))
thetas <- rstan::extract(fit, "theta")[[1]]

### get rank information
### over iterations, get rank
ranks <- apply(thetas, 1, rank)
### Summarize these ranks
ranks <- apply(ranks, 1, function(x) {
    data.frame(mean_rank = round(mean(x)),
               lo_rank = round(quantile(x, 0.025)),
               hi_rank = round(quantile(x, 1 - 0.025)))
})
ranks <- bind_rows(ranks) |>
    mutate(mp_idx = 1:n())

plot_df <- left_join(plot_df, ranks, by = join_by(mp_idx))
plot_df <- plot_df |>
    arrange(mean) |>
    mutate(actual_rank = 1:n())


tab <- plot_df |>
    arrange(mean) |>
    mutate(DisplayName = sub(" and ", " & ", DisplayName)) |>
    mutate(DisplayName = sub(" \\(.*", "", DisplayName)) |>
    mutate(Party = dplyr::recode(Party,
                                 "Scottish National Party" = "SNP",
                                 "Conservative" = "Cons.")) |>
    dplyr::select(MP = DisplayName, Party,
                  `Mean` = mean, `Lo` = `2.5%`, `Hi` = `97.5%`,
                  `Rank` = actual_rank, `Lo ` = `lo_rank`, `Hi ` = `hi_rank`)

kbl(tab,
    digits = 2,
    format = "latex",
    booktabs = TRUE,
    longtable = TRUE) |>
    add_header_above(c(" " = 2, "Position" = 3, "Rank" = 3)) |>
    kable_styling(latex_options = c("repeat_header"))

```

{{< pagebreak >}}

# Stan code

```{r}
writeLines(readLines(here::here("working", "partyonly.stan")))
```

# Efficiency gains over dichotomous or trichotomous responses

```{r trchot}
trichot_df <- readRDS(here::here("working", "trichot_thetas.rds")) |>
    mutate(Party = factor(Party, ordered = TRUE),
           Party = reorder(Party, mean)) 

ci_width <- plot_df |>
    mutate(ci_width = `97.5%` - `2.5%`) |>
    summarize(ci_width = mean(ci_width)) |>
    pull(ci_width) |>
    round(2)


ci_width_trichot <- trichot_df |>
    mutate(ci_width = `97.5%` - `2.5%`) |>
    summarize(ci_width = mean(ci_width)) |>
    pull(ci_width) |>
    round(2)

joint <- left_join(plot_df, trichot_df,
                   by = join_by(DisplayName, Party, PCON22CD, Person.ID))

the_cor <- with(joint,
                cor(mean.x, mean.y))
the_cor <- round(the_cor, 2)
```

The average CI width in the five-category model was `r ci_width`. 
This compares to a SD of `r round(sd(plot_df$mean), 2)`. The CI width expressed as a proportion of the SD is therefore `r round(ci_width / sd(plot_df$mean), 2)`. 

The average CI width in the three-category model was `r ci_width_trichot`. 
This compares to a SD of `r round(sd(trichot_df$mean), 2)`. The CI width expressed as a proportion of the SD is therefore `r round(ci_width_trichot / sd(trichot_df$mean), 2)`. 

The correlation between the two measures is $r=`r the_cor`$.
